---
title: "Construct County CESI Scores"
author: "Erin Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: readable
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))

knitr::opts_chunk$set(
  echo = TRUE,
  warning = TRUE,
  message = FALSE,
  error = FALSE)

library(dplyr)
library(tidyr)
library(readr)
library(stringr)
```
  
This script is for creating teh two social interdependence indexes.

# File paths

```{r paths}
# Keep the source data separate from the derived CESI files.
county_data <- "data/raw/county_collectivism"
output_dir <- "data/derived/cesi_scores"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
```

# Practice-based CESI

Practice-based CESI measures how social life is organized through observable county practices (following V&C 1999).

## Libertarian vote share

The voting indicator is the average county Libertarian vote share across the 2008, 2012, 2016, and 2020 presidential elections. Some counties are missing an election year, so the average uses the available years.

```{r libertarian-vote}
# Read each election separately because the source files use year-specific variable names
libertarian_2008 <- read_csv(
  file.path(county_data, "libertarian", "libertarian08.csv"),
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE) %>%
  transmute(
    county_fips = str_pad(county_fips, 5, pad = "0"),
    libertarian_2008 = libertarian_08)

libertarian_2012 <- read_csv(
  file.path(county_data, "libertarian", "libertarian12.csv"),
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE) %>%
  transmute(
    county_fips = str_pad(county_fips, 5, pad = "0"),
    libertarian_2012 = libertarian_12)

libertarian_2016 <- read_csv(
  file.path(county_data, "libertarian", "libertarian16.csv"),
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE) %>%
  transmute(
    county_fips = str_pad(county_fips, 5, pad = "0"),
    libertarian_2016 = libertarian_16)

# The 2020 file has multiple ballot-mode rows within a county
libertarian_2020 <- suppressMessages(read_csv(
  file.path(county_data, "libertarian", "libertarian20_all.csv"),
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE)) %>%
  mutate(county_fips = str_pad(county_fips, 5, pad = "0")) %>%
  group_by(county_fips) %>%
  summarise(
    libertarian_votes = sum(candidatevotes),
    total_votes = max(totalvotes),
    .groups = "drop") %>%
  mutate(libertarian_2020 = libertarian_votes / total_votes) %>%
  select(county_fips, libertarian_2020)

# Average across available elections
libertarian_vote <- libertarian_2008 %>%
  left_join(libertarian_2012, by = "county_fips") %>%
  left_join(libertarian_2016, by = "county_fips") %>%
  left_join(libertarian_2020, by = "county_fips") %>%
  mutate(
    n_elections = rowSums(!is.na(pick(starts_with("libertarian_20")))),
    libertarian_vote = rowMeans(
      pick(starts_with("libertarian_20")),
      na.rm = TRUE),
    libertarian_vote = if_else(
      n_elections == 0,
      NA_real_,
      libertarian_vote)) %>%
  select(county_fips, libertarian_vote, n_elections)

# Check for duplicates
libertarian_vote %>%
  count(county_fips) %>%
  filter(n > 1)

```

## ACS and religious-adherence 

The ACS file contains six pre-calculated county proportions or ratios: self-employment, marriage relative to divorce, living alone, older adults living alone, multigenerational living, and carpooling relative to driving alone. Religious adherence is the number of reported adherents divided by county population in the ARDA 2020 data.

```{r practice-inputs}
# Read the county practice indicators and standardize the FIPS identifier
acs <- read_csv(
  file.path(
    county_data,
    "acs",
    "ACS22_5year_county.csv"),
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE,
  name_repair = "minimal") %>%
  separate(county_state_name, into = c("county", "state"), sep = ", ") %>%
  mutate(
    county_fips = str_pad(county_fips, 5, pad = "0"),
    county = str_remove(county, "\\s+County$"),
    county = str_remove(county, " Parish$"),
    county = str_replace(county, " city$", " City")) %>%
  select(
    county_fips,
    county,
    state,
    self_employed,
    marriage_divorce_rate,
    living_alone,
    living_alone_65over,
    living_grandchildren,
    carpool_drive_alone)

religion <- read_csv(
  file.path(county_data, "religion", "adherence_rate_county.csv"),
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE,
  name_repair = "minimal") %>%
  transmute(
    county_fips = str_pad(county_fips, 5, pad = "0"),
    religious_adherents) %>%
  filter(str_detect(county_fips, "^[0-9]{5}$"))

# Check for duplicates
bind_rows(
  acs %>% count(county_fips) %>% mutate(dataset = "ACS"),
  religion %>% count(county_fips) %>% mutate(dataset = "Religion")) %>%
  filter(n > 1)

# Merge
practice_data <- acs %>%
  left_join(religion, by = "county_fips") %>%
  left_join(libertarian_vote, by = "county_fips")

```

## Construct the practice score

```{r practice-score}
# Check descriptives (religious adherence and the divorce-to-marriage ratio are skewed)
practice_data <- practice_data %>%
  mutate(
    divorce_marriage_rate = 1 / marriage_divorce_rate,
    divorce_marriage_rate_log = log(divorce_marriage_rate),
    religious_adherents_log = if_else(
      religious_adherents > 0,
      log(religious_adherents),
      NA_real_))

transformation_check <- practice_data %>%
  select(
    religious_adherents,
    religious_adherents_log,
    divorce_marriage_rate,
    divorce_marriage_rate_log) %>%
  psych::describe() %>%
  as.data.frame() %>%
  tibble::rownames_to_column("variable") %>%
  select(variable, n, mean, sd, median, min, max, skew, kurtosis)

transformation_check 

# Standardize the indicators
practice_scores <- practice_data %>%
  mutate(
    carpool_interdependence_z =
      as.numeric(scale(carpool_drive_alone)),
    multigenerational_interdependence_z =
      as.numeric(scale(living_grandchildren)),
    religious_interdependence_raw_z =
      as.numeric(scale(religious_adherents)),
    religious_interdependence_log_z =
      as.numeric(scale(religious_adherents_log)),
    living_alone_independence_z =
      as.numeric(scale(living_alone)),
    elder_living_alone_independence_z =
      as.numeric(scale(living_alone_65over)),
    divorce_marriage_independence_raw_z =
      as.numeric(scale(divorce_marriage_rate)),
    divorce_marriage_independence_log_z =
      as.numeric(scale(divorce_marriage_rate_log)),
    libertarian_independence_z =
      as.numeric(scale(libertarian_vote)),
    self_employed_independence_z =
      as.numeric(scale(self_employed))) %>%
  mutate(
    interdependence_score_raw = rowMeans(pick(
      carpool_interdependence_z,
      multigenerational_interdependence_z,
      religious_interdependence_raw_z)),
    independence_score_raw = rowMeans(pick(
      living_alone_independence_z,
      elder_living_alone_independence_z,
      divorce_marriage_independence_raw_z,
      libertarian_independence_z,
      self_employed_independence_z)),
    interdependence_score_log = rowMeans(pick(
      carpool_interdependence_z,
      multigenerational_interdependence_z,
      religious_interdependence_log_z)),
    independence_score_log = rowMeans(pick(
      living_alone_independence_z,
      elder_living_alone_independence_z,
      divorce_marriage_independence_log_z,
      libertarian_independence_z,
      self_employed_independence_z))) %>%
  mutate(
    interdependence_score_raw_z =
      as.numeric(scale(interdependence_score_raw)),
    independence_score_raw_z =
      as.numeric(scale(independence_score_raw)),
    practice_cesi_raw =
      interdependence_score_raw_z - independence_score_raw_z,
    interdependence_score_log_z =
      as.numeric(scale(interdependence_score_log)),
    independence_score_log_z =
      as.numeric(scale(independence_score_log)),
    practice_cesi =
      interdependence_score_log_z - independence_score_log_z)

cat("Counties with Practice CESI:", sum(!is.na(practice_scores$practice_cesi)), "\n") #3100

```

# Meaning-based CESI

Havaldar et al. (2024) developed county scores for collectivist and individualist language. 
- Collectivism lexicon: duties, responsibilities, role, fit in, community, sacrifice, shame, required, rules, honor, support, rely, loyal, respect, obedience.
- Individualism lexicon: human, humanity, worldwide, universal, mankind, everyone, collective, global, equity, imagination, cooperative, cooperation, shared, joint, identity, guilty, diversity.

```{r meaning-score}
# Read and standardize 
meaning_scores <- read_csv(
  "data/raw/meaning_cesi/county_scores.csv",
  col_types = cols(group_id = col_character()),
  show_col_types = FALSE) %>%
  filter(
    group_id != "0",
    feat %in% c("COLLECTIVIST", "INDIVIDUALIST")) %>%
  select(group_id, feat, group_norm) %>%
  pivot_wider(names_from = feat, values_from = group_norm) %>%
  filter(!is.na(COLLECTIVIST), !is.na(INDIVIDUALIST)) %>%
  mutate(
    county_fips = str_pad(group_id, 5, pad = "0"),
    collectivist_language_z = as.numeric(scale(COLLECTIVIST)),
    individualist_language_z = as.numeric(scale(INDIVIDUALIST)),
    meaning_cesi_unstandardized =
      collectivist_language_z - individualist_language_z,
    meaning_cesi = as.numeric(scale(meaning_cesi_unstandardized))) %>%
  select(
    county_fips,
    collectivist_language = COLLECTIVIST,
    individualist_language = INDIVIDUALIST,
    collectivist_language_z,
    individualist_language_z,
    meaning_cesi_unstandardized,
    meaning_cesi)

# Check for duplicates
meaning_scores %>%
  count(county_fips) %>%
  filter(n > 1)

cat("Counties with Meaning CESI:", nrow(meaning_scores), "\n") #2041

```

# Combine the two measures

```{r combine-scores}
# Join files
county_cesi_scores <- practice_scores %>%
  select(
    county_fips,
    county,
    state,
    n_elections,
    interdependence_score_raw,
    independence_score_raw,
    practice_cesi_raw,
    interdependence_score_log,
    independence_score_log,
    practice_cesi) %>%
  full_join(meaning_scores, by = "county_fips") %>%
  arrange(county_fips)

# Check for duplicates
county_cesi_scores %>%
  count(county_fips) %>%
  filter(n > 1)

# Check # counties for each index
score_coverage <- county_cesi_scores %>%
  summarise(
    total_counties = n(),
    practice_cesi_n = sum(!is.na(practice_cesi)),
    meaning_cesi_n = sum(!is.na(meaning_cesi)),
    both_scores_n = sum(!is.na(practice_cesi) & !is.na(meaning_cesi)))

score_coverage

```

# Export

```{r export}
write_csv(county_cesi_scores, file.path(output_dir, "county_cesi_scores.csv"))
write_csv(score_coverage, file.path(output_dir, "cesi_score_coverage.csv"))
```
