---
title: "Combined County Health Outcome Dataset"
subtitle: "From CDC PLACES and County Health Rankings & Roadmaps"
author: "Erin Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: readable
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE
)

library(dplyr)
library(tidyr)
library(readr)
library(stringr)
```

This script is for cleaning and merging health datsets from CDC PLACES Aand CHR&R.

# File paths

```{r paths}
# Input
places_path <- "data/raw/cdc_places_2024/PLACES__Local_Data_for_Better_Health,_County_Data_2024_release_20260710.csv"
chr_path <- "data/raw/county_health_rankings_2025_2026/analytic_data2025_v3.csv"

# Output
derived_dir <- "data/derived/combined_health_outcomes"

dir.create(derived_dir, recursive = TRUE, showWarnings = FALSE)

```

# CDC PLACES

## Data exclusions

The existing PLACES file includes health outcomes, health behaviors, disabilities, social needs, and healthcare-use measures. I kept measures of health status, chronic conditions, disability, and health-risk behavior. I excluded the Health-Related Social Needs and Prevention categories because those variables are more appropriately treated as predictors or resources than as health outcomes. This removes food insecurity, food stamps, housing insecurity, transportation barriers, utility shut-off threat, insurance, medication use, checkups, screening, dental visits, and mammography. I retained social isolation (`ISOLATION`) and lack of social or emotional support (`EMOTIONSPT`) as psychosocial outcomes relevant to social interdependence.

```{r places-exclusion-rule}
# Define exclusions
excluded_places_categories <- c(
  "Health-Related Social Needs",
  "Prevention")

included_places_exceptions <- c(
  "ISOLATION",
  "EMOTIONSPT")

expected_excluded_places_ids <- c(
  "FOODINSECU",
  "FOODSTAMP",
  "HOUSINSECU",
  "LACKTRPT",
  "SHUTUTILITY",
  "ACCESS2",
  "BPMED",
  "CHECKUP",
  "CHOLSCREEN",
  "COLON_SCREEN",
  "DENTAL",
  "MAMMOUSE")

```

# Read in data

```{r read-places}
# Read county-level PLACES data
places_raw <- read_csv(
  places_path,
  col_types = cols(
    .default = col_character(),
    LocationID = col_character(),
    Data_Value = col_character(),
    Low_Confidence_Limit = col_character(),
    High_Confidence_Limit = col_character()),
  show_col_types = FALSE)

available_measure_ids <- unique(places_raw$MeasureId)
setdiff(expected_excluded_places_ids, available_measure_ids)

# Use age-adjusted county estimates
places_health_long <- places_raw %>%
  filter(
    Data_Value_Type == "Age-adjusted prevalence",
    StateAbbr != "US",
    !is.na(Data_Value),
    !is.na(LocationName),
    !Category %in% excluded_places_categories |
      MeasureId %in% included_places_exceptions) %>%
  transmute(
    county_fips = str_pad(LocationID, 5, pad = "0"),
    state = StateAbbr,
    county = LocationName,
    source = "CDC PLACES 2024",
    source_release = "2024 release",
    outcome_id = MeasureId,
    outcome_label = Short_Question_Text,
    outcome_family = Category,
    outcome_value = parse_number(Data_Value),
    native_unit = "Age-adjusted percentage of adults",
    higher_value_interpretation = "Worse health / higher risk prevalence",
    definition = Measure,
    ci_low = parse_number(Low_Confidence_Limit),
    ci_high = parse_number(High_Confidence_Limit)) %>%
  filter(str_detect(county_fips, "^[0-9]{5}$"))

# Check duplicates
places_duplicates <- places_health_long %>%
  count(county_fips, outcome_id) %>%
  filter(n > 1)

```

# CHR&R

County Health Rankings & Roadmaps has additional data beyond CDC PLACES, which capture longevity, mortality, birth outcomes, and infectious disease and do not
duplicate the current PLACES variables. Duplicate/similar outcome variables (such as poor general health, mental distress, physical distress, diabetes) are not included.

```{r chr-outcome-dictionary}
# List the CHR&R outcomes
chr_outcome_dictionary <- tribble(
  ~chr_column, ~outcome_id, ~outcome_label, ~outcome_family, ~native_unit,
  ~higher_value_interpretation, ~definition,
  "Life Expectancy raw value", "CHR_LIFEEXP", "Life Expectancy",
  "Mortality and longevity", "Years", "Better health / greater longevity",
  "Average number of years people are expected to live.",
  "Premature Death raw value", "CHR_PREMATURE_DEATH", "Premature Death",
  "Mortality and longevity", "Years of potential life lost before age 75 per 100,000 (age-adjusted)",
  "Worse health / higher mortality burden",
  "Years of potential life lost before age 75 per 100,000 population (age-adjusted).",
  "Premature Age-Adjusted Mortality raw value", "CHR_PREMATURE_AAM", "Premature Age-Adjusted Mortality",
  "Mortality and longevity", "Deaths per 100,000 (age-adjusted)",
  "Worse health / higher mortality",
  "Deaths among residents under age 75 per 100,000 population (age-adjusted).",
  "Child Mortality raw value", "CHR_CHILD_MORTALITY", "Child Mortality",
  "Mortality and longevity", "Deaths under age 20 per 100,000",
  "Worse health / higher mortality",
  "Deaths among residents under age 20 per 100,000 population.",
  "Infant Mortality raw value", "CHR_INFANT_MORTALITY", "Infant Mortality",
  "Mortality and longevity", "Deaths under age 1 per 1,000 live births",
  "Worse health / higher mortality",
  "Infant deaths within the first year of life per 1,000 live births.",
  "Injury Deaths raw value", "CHR_INJURY_DEATHS", "Injury Deaths",
  "Mortality and longevity", "Deaths per 100,000",
  "Worse health / higher mortality",
  "Deaths due to injury per 100,000 population.",
  "Suicides raw value", "CHR_SUICIDES", "Suicide Mortality",
  "Mortality and longevity", "Deaths per 100,000 (age-adjusted)",
  "Worse health / higher mortality",
  "Deaths due to suicide per 100,000 population (age-adjusted).",
  "Drug Overdose Deaths raw value", "CHR_OVERDOSE_DEATHS", "Drug Overdose Deaths",
  "Mortality and longevity", "Deaths per 100,000",
  "Worse health / higher mortality",
  "Drug poisoning deaths per 100,000 population.",
  "Homicides raw value", "CHR_HOMICIDES", "Homicide Mortality",
  "Mortality and longevity", "Deaths per 100,000",
  "Worse health / higher mortality",
  "Deaths due to homicide per 100,000 population.",
  "Firearm Fatalities raw value", "CHR_FIREARM_FATALITIES", "Firearm Fatalities",
  "Mortality and longevity", "Deaths per 100,000",
  "Worse health / higher mortality",
  "Deaths due to firearms per 100,000 population.",
  "Motor Vehicle Crash Deaths raw value", "CHR_MVC_DEATHS", "Motor Vehicle Crash Deaths",
  "Mortality and longevity", "Deaths per 100,000",
  "Worse health / higher mortality",
  "Motor vehicle crash deaths per 100,000 population.",
  "Alcohol-Impaired Driving Deaths raw value", "CHR_ALCOHOL_DRIVING_DEATHS", "Alcohol-Impaired Driving Deaths",
  "Mortality and longevity", "Proportion of driving deaths involving alcohol impairment",
  "Worse health / greater alcohol-related mortality burden",
  "Proportion of driving deaths involving alcohol-impaired drivers.",
  "Low Birth Weight raw value", "CHR_LOW_BIRTH_WEIGHT", "Low Birth Weight",
  "Reproductive and infectious health", "Proportion of live births",
  "Worse population health",
  "Live births with birth weight below 2,500 grams.",
  "Teen Births raw value", "CHR_TEEN_BIRTHS", "Teen Births",
  "Reproductive and infectious health", "Births per 1,000 females ages 15-19",
  "Higher reproductive-health risk",
  "Births per 1,000 female population ages 15-19.",
  "HIV Prevalence raw value", "CHR_HIV_PREVALENCE", "HIV Prevalence",
  "Reproductive and infectious health", "People living with diagnosed HIV per 100,000 age 13+",
  "Higher infectious-disease burden",
  "People age 13 and older living with diagnosed HIV infection per 100,000 population.",
  "Sexually Transmitted Infections raw value", "CHR_STI", "Sexually Transmitted Infections",
  "Reproductive and infectious health", "Diagnosed cases per 100,000",
  "Higher infectious-disease burden",
  "Diagnosed sexually transmitted infections per 100,000 population."
)
```

## Read and reshape CHR&R

```{r read-chr}
# Read in data
chr_raw <- read_csv(chr_path, show_col_types = FALSE)

# Remove national and non-county rows, reshape outcomes to long form, add definitions
chr_health_long <- chr_raw %>%
  transmute(
    county_fips = as.character(`5-digit FIPS Code`),
    state = `State Abbreviation`,
    county = Name,
    across(all_of(chr_outcome_dictionary$chr_column))) %>%
  filter(
    str_detect(county_fips, "^[0-9]{5}$"),
    state != "US",
    !is.na(county),
    str_trim(county) != "",
    str_to_lower(str_trim(county)) != "united states") %>%
  distinct(county_fips, .keep_all = TRUE) %>%
  pivot_longer(
    cols = all_of(chr_outcome_dictionary$chr_column),
    names_to = "chr_column",
    values_to = "source_value") %>%
  mutate(outcome_value = parse_number(as.character(source_value))) %>%
  select(-source_value) %>%
  left_join(chr_outcome_dictionary, by = "chr_column") %>%
  transmute(
    county_fips,
    state,
    county,
    source = "County Health Rankings & Roadmaps 2025",
    source_release = "2025 analytic data v3",
    outcome_id,
    outcome_label,
    outcome_family,
    outcome_value,
    native_unit,
    higher_value_interpretation,
    definition,
    ci_low = NA_real_,
    ci_high = NA_real_) %>%
  filter(!is.na(outcome_value))

```

# Combine

```{r combine-and-validate}
# Combine two datasets
combined_health_long <- bind_rows(
  places_health_long,
  chr_health_long) %>%
  arrange(source, outcome_family, outcome_id, county_fips)

# Check duplicates
combined_duplicates <- combined_health_long %>%
  count(county_fips, outcome_id) %>%
  filter(n > 1)
combined_duplicates

county_state_duplicates <- combined_health_long %>%
  distinct(county_fips, state) %>%
  count(county_fips) %>%
  filter(n > 1)
county_state_duplicates

# County labels differ slightly across sources, so use the PLACES label when it is available
county_lookup <- bind_rows(
  places_health_long %>%
    distinct(county_fips, state, county) %>%
    mutate(label_priority = 1L),
  chr_health_long %>%
    distinct(county_fips, state, county) %>%
    mutate(label_priority = 2L)) %>%
  arrange(county_fips, label_priority) %>%
  distinct(county_fips, .keep_all = TRUE) %>%
  select(-label_priority)

# Create a wide dataset
combined_health_wide <- combined_health_long %>%
  select(county_fips, outcome_id, outcome_value) %>%
  pivot_wider(names_from = outcome_id, values_from = outcome_value) %>%
  left_join(county_lookup, by = "county_fips") %>%
  relocate(county_fips, state, county)

# Create a variables dictionary
combined_dictionary <- combined_health_long %>%
  distinct(
    source,
    source_release,
    outcome_id,
    outcome_label,
    outcome_family,
    native_unit,
    higher_value_interpretation,
    definition) %>%
  arrange(source, outcome_family, outcome_id)

```

# Export

```{r write-files}
write_csv(combined_health_long, file.path(derived_dir, "combined_direct_health_outcomes_long.csv"))
write_csv(combined_health_wide, file.path(derived_dir, "combined_direct_health_outcomes_wide.csv"))
write_csv(combined_dictionary, file.path(derived_dir, "combined_direct_health_outcome_dictionary.csv"))

```
