---
title: "Construct County Adversity Score"
author: "Erin Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: readable
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))

knitr::opts_chunk$set(
  echo = TRUE,
  warning = TRUE,
  message = FALSE,
  error = FALSE)

library(dplyr)
library(tidyr)
library(readr)
library(stringr)
```

# File paths

```{r paths}
output_dir <- "data/derived/adversity_index"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

```

# Socioeconomic disadvantage

The two ACS indicators are low household income and low educational attainment.

```{r acs}
# Reverse the income and education measures so higher values consistently mean greater adversity; log income - right-skewed
acs <- read_csv(
  "data/raw/acs_2022_5year/acs2022_5year_county_income_education.csv",
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE) %>%
  transmute(
    county_fips = str_pad(county_fips, 5, pad = "0"),
    low_income = -log(replace(
      median_household_income_2022,
      median_household_income_2022 <= 0,
      NA_real_)),
    low_education = 100 - bachelors_or_higher_pct)

# A blank result confirms one ACS row per county.
acs %>%
  count(county_fips) %>%
  filter(n > 1)

```

# Material hardship

There are five age-adjusted PLACES measures included: food insecurity, use of food stamps,
housing insecurity, transportation barriers, and threat of utility shutoff.

```{r places}
# Read in data
places <- read_csv(
  "data/raw/cdc_places_2024/PLACES__Local_Data_for_Better_Health,_County_Data_2024_release_20260710.csv",
  col_types = cols(
    .default = col_character(),
    LocationID = col_character(),
    Data_Value = col_character()),
  show_col_types = FALSE) %>%
  filter(
    Data_Value_Type == "Age-adjusted prevalence",
    MeasureId %in% c(
      "FOODINSECU", "FOODSTAMP", "HOUSINSECU", "LACKTRPT", "SHUTUTILITY"),
    StateAbbr != "US") %>%
  transmute(
    county_fips = str_pad(LocationID, 5, pad = "0"),
    measure_id = MeasureId,
    value = parse_number(Data_Value)) %>%
  distinct(county_fips, measure_id, .keep_all = TRUE) %>%
  pivot_wider(names_from = measure_id, values_from = value) %>%
  rename(
    food_insecurity = FOODINSECU,
    food_stamps = FOODSTAMP,
    housing_insecurity = HOUSINSECU,
    transportation_barriers = LACKTRPT,
    utility_threat = SHUTUTILITY)

# Check duplicates
places %>%
  count(county_fips) %>%
  filter(n > 1)

```

# Healthcare access and preventive resources

```{r county-health-rankings}
# Read in data
chr_2025 <- read_csv(
  "data/raw/county_health_rankings_2025_2026/analytic_data2025_v3.csv",
  show_col_types = FALSE) %>%
  mutate(
    county_fips = as.character(`5-digit FIPS Code`),
    state = `State Abbreviation`,
    county = Name) %>%
  filter(
    str_detect(county_fips, "^[0-9]{5}$"),
    state != "US",
    !is.na(county),
    county != "United States") %>%
  transmute(
    county_fips,
    state,
    county,
    flu_2025 = parse_number(as.character(`Flu Vaccinations raw value`)),
    exercise_2025 = parse_number(as.character(`Access to Exercise Opportunities raw value`)),
    food_environment_2025 = parse_number(as.character(`Food Environment Index raw value`)),
    primary_care_2025 = parse_number(as.character(`Primary Care Physicians raw value`)),
    mental_health_2025 = parse_number(as.character(`Mental Health Providers raw value`)),
    dentists_2025 = parse_number(as.character(`Dentists raw value`)),
    mammography_2025 = parse_number(as.character(`Mammography Screening raw value`)),
    uninsured_2025 = parse_number(as.character(`Uninsured raw value`)),
    life_expectancy = parse_number(as.character(`Life Expectancy raw value`)))

# Read the supplement data
chr_2026 <- read_csv(
  "data/raw/county_health_rankings_2025_2026/analytic_supplement_20260325.csv",
  col_types = cols(.default = col_character()),
  show_col_types = FALSE) %>%
  mutate(county_fips = str_pad(fipscode, 5, pad = "0")) %>%
  filter(str_detect(county_fips, "^[0-9]{5}$")) %>%
  transmute(
    county_fips,
    flu_2026 = parse_number(v155_rawvalue),
    food_environment_2026 = parse_number(v133_rawvalue),
    primary_care_2026 = parse_number(v004_rawvalue),
    mental_health_2026 = parse_number(v062_rawvalue),
    dentists_2026 = parse_number(v088_rawvalue),
    mammography_2026 = parse_number(v050_rawvalue),
    uninsured_2026 = parse_number(v085_rawvalue))

# Reverse code so every measure points toward adversity
chr <- chr_2025 %>%
  full_join(chr_2026, by = "county_fips") %>%
  transmute(
    county_fips,
    state,
    county,
    low_flu_vaccination = -coalesce(flu_2026, flu_2025),
    low_exercise_access = -exercise_2025,
    poor_food_environment =
      -coalesce(food_environment_2026, food_environment_2025),
    few_primary_care_physicians =
      -log1p(coalesce(primary_care_2026, primary_care_2025) * 100000),
    few_mental_health_providers =
      -log1p(coalesce(mental_health_2026, mental_health_2025) * 100000),
    few_dentists =
      -log1p(coalesce(dentists_2026, dentists_2025) * 100000),
    low_mammography_screening =
      -coalesce(mammography_2026, mammography_2025),
    uninsured = coalesce(uninsured_2026, uninsured_2025),
    life_expectancy)

# Check duplicates
chr %>%
  count(county_fips) %>%
  filter(n > 1)

```

# Combine indicators

```{r combine}
# Merge
adversity_data <- chr %>%
  full_join(acs, by = "county_fips") %>%
  full_join(places, by = "county_fips") %>%
  filter(
    str_detect(county_fips, "^[0-9]{5}$"),
    state != "US",
    !is.na(county),
    str_trim(county) != "",
    str_to_lower(str_trim(county)) != "united states") %>%
  distinct(county_fips, .keep_all = TRUE)

adversity_indicators <- c(
  "low_income",
  "low_education",
  "food_insecurity",
  "food_stamps",
  "housing_insecurity",
  "transportation_barriers",
  "utility_threat",
  "few_primary_care_physicians",
  "few_mental_health_providers",
  "few_dentists",
  "uninsured",
  "low_flu_vaccination",
  "low_exercise_access",
  "poor_food_environment",
  "low_mammography_screening")

# Check missingness
indicator_missingness <- adversity_data %>%
  summarise(across(all_of(adversity_indicators), ~ sum(is.na(.x)))) %>%
  pivot_longer(
    everything(),
    names_to = "indicator",
    values_to = "n_missing") %>%
  mutate(percent_missing = 100 * n_missing / nrow(adversity_data))
indicator_missingness

adversity_standardized <- adversity_data

# Winsorize each indicator
for (variable in adversity_indicators) {
  limits <- quantile(
    adversity_standardized[[variable]],
    c(.01, .99),
    na.rm = TRUE)

  adversity_standardized[[variable]] <- pmin(
    pmax(adversity_standardized[[variable]], limits[1]),
    limits[2])

  adversity_standardized[[variable]] <- as.numeric(
    scale(adversity_standardized[[variable]]))
}

```

# Construct the adversity score

A county must have all 15 indicators. Indicators receive equal weight within
each domain, and the four standardized domains receive equal weight in the
overall score.

```{r adversity-score}
adversity_scores <- adversity_standardized %>%
  drop_na(all_of(adversity_indicators)) %>%
  transmute(
    county_fips,
    state,
    county,
    socioeconomic = rowMeans(pick(low_income, low_education)),
    material_hardship = rowMeans(pick(
      food_insecurity,
      food_stamps,
      housing_insecurity,
      transportation_barriers,
      utility_threat)),
    healthcare_access = rowMeans(pick(
      few_primary_care_physicians,
      few_mental_health_providers,
      few_dentists,
      uninsured)),
    preventive_resources = rowMeans(pick(
      low_flu_vaccination,
      low_exercise_access,
      poor_food_environment,
      low_mammography_screening)),
    life_expectancy) %>%
  mutate(
    socioeconomic_z = as.numeric(scale(socioeconomic)),
    material_hardship_z = as.numeric(scale(material_hardship)),
    healthcare_access_z = as.numeric(scale(healthcare_access)),
    preventive_resources_z = as.numeric(scale(preventive_resources)),
    adversity_domain_balanced = as.numeric(scale(rowMeans(pick(
      socioeconomic_z,
      material_hardship_z,
      healthcare_access_z,
      preventive_resources_z)))))

cat("Counties with an adversity score:", nrow(adversity_scores), "\n")

# Correctional check with Life expectancy as a quick sanity check
adversity_scores %>%
  summarise(
    correlation_with_life_expectancy = cor(
      adversity_domain_balanced,
      life_expectancy,
      use = "complete.obs"))

```

# Variable dictionary

```{r dictionary}
adversity_dictionary <- tribble(
  ~domain, ~indicator, ~interpretation,
  "Socioeconomic disadvantage", "low_income", "Lower median household income",
  "Socioeconomic disadvantage", "low_education", "Lower percentage with a bachelor's degree",
  "Material hardship", "food_insecurity", "Higher prevalence of food insecurity",
  "Material hardship", "food_stamps", "Higher prevalence of food-stamp use",
  "Material hardship", "housing_insecurity", "Higher prevalence of housing insecurity",
  "Material hardship", "transportation_barriers", "Higher prevalence of transportation barriers",
  "Material hardship", "utility_threat", "Higher prevalence of utility shutoff threat",
  "Healthcare access", "few_primary_care_physicians", "Fewer primary-care physicians",
  "Healthcare access", "few_mental_health_providers", "Fewer mental-health providers",
  "Healthcare access", "few_dentists", "Fewer dentists",
  "Healthcare access", "uninsured", "Higher uninsured percentage",
  "Preventive resources", "low_flu_vaccination", "Lower flu-vaccination coverage",
  "Preventive resources", "low_exercise_access", "Lower access to exercise opportunities",
  "Preventive resources", "poor_food_environment", "Lower food-environment score",
  "Preventive resources", "low_mammography_screening", "Lower mammography-screening coverage")
adversity_dictionary

```

# Export

```{r export}
write_csv(adversity_scores, file.path(output_dir, "county_adversity_domain_balanced_scores.csv"))
write_csv( adversity_dictionary, file.path(output_dir, "adversity_score_dictionary.csv"))
write_csv(indicator_missingness, file.path(output_dir, "adversity_indicator_missingness.csv"))

```
