---
title: "Primary CESI and County Health Models"
author: "Erin Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: readable
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))

knitr::opts_chunk$set(
  echo = TRUE,
  warning = TRUE,
  message = FALSE,
  error = FALSE)

library(dplyr)
library(tidyr)
library(readr)
library(stringr)
```

This script runs the primary county-health analyses. Counties are nested within states, so both models include a state random intercept. 

# Read and combine the data

```{r read-data}
# Read two final CESI scores
cesi <- read_csv(
  "data/derived/cesi_scores/county_cesi_scores.csv",
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE) %>%
  select(county_fips, practice_cesi, meaning_cesi)

# Read in data
health <- read_csv(
  "data/derived/combined_health_outcomes/combined_direct_health_outcomes_long.csv",
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE) %>%
  transmute(
    county_fips = str_pad(county_fips, 5, pad = "0"),
    state = state,
    source,
    outcome_id,
    outcome = outcome_label,
    outcome_value,
    native_unit,
    higher_value_interpretation)

adversity <- read_csv(
  "data/derived/adversity_index/county_adversity_domain_balanced_scores.csv",
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE) %>%
  select(
    county_fips,
    adversity_domain_balanced,
    socioeconomic_z,
    material_hardship_z,
    healthcare_access_z,
    preventive_resources_z)

demographics <- read_csv(
  "data/raw/covariates/county_demographics_2022.csv",
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE) %>%
  transmute(
    county_fips = str_pad(county_fips, 5, pad = "0"),
    county_age = county_age22)

race <- read_csv(
  "data/raw/covariates/county-race.csv",
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE) %>%
  transmute(
    county_fips = str_pad(county_fips, 5, pad = "0"),
    percent_minority = 1 - pct_white)

rurality <- suppressMessages(read_csv(
  "data/raw/covariates/county_rural_score.csv",
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE)) %>%
  transmute(
    county_fips = str_pad(county_fips, 5, pad = "0"),
    rural_urban_code = as.numeric(RUCC_2023))

density <- read_csv(
  "data/raw/covariates/county_income.csv",
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE) %>%
  transmute(
    county_fips = str_pad(county_fips, 5, pad = "0"),
    log_population_density = log1p(pop_density))

# Join data
health_data <- health %>%
  left_join(cesi, by = "county_fips") %>%
  left_join(adversity, by = "county_fips") %>%
  left_join(demographics, by = "county_fips") %>%
  left_join(race, by = "county_fips") %>%
  left_join(rurality, by = "county_fips") %>%
  left_join(density, by = "county_fips")

# Check duplicates
health_data %>%
  count(county_fips, outcome_id) %>%
  filter(n > 1)

```

# Sample coverage

```{r sample-coverage}
sample_coverage <- health_data %>%
  group_by(outcome_id, outcome) %>%
  summarise(
    outcome_n = sum(!is.na(outcome_value)),
    practice_n = sum(!is.na(outcome_value) & !is.na(practice_cesi)),
    meaning_n = sum(!is.na(outcome_value) & !is.na(meaning_cesi)),
    practice_m2_n = sum(complete.cases(
      outcome_value,
      practice_cesi,
      adversity_domain_balanced,
      log_population_density,
      rural_urban_code,
      county_age,
      percent_minority)),
    meaning_m2_n = sum(complete.cases(
      outcome_value,
      meaning_cesi,
      adversity_domain_balanced,
      log_population_density,
      rural_urban_code,
      county_age,
      percent_minority)),
    .groups = "drop")

sample_coverage

```

# Primary models

M1 adjusts for basic county characteristics. M2 adds the domain-balanced adversity score. M1 and M2 are fit on the same complete-case sample for each outcome and CESI measure for comparison (adjust for adveristy).

```{r model-function}
run_primary_models <- function(data, selected_outcome, cesi_variable) {
  model_data <- data %>%
    filter(outcome_id == selected_outcome) %>%
    transmute(
      state = factor(state),
      health = outcome_value,
      cesi = .data[[cesi_variable]],
      adversity = adversity_domain_balanced,
      population_density = log_population_density,
      rurality = rural_urban_code,
      age = county_age,
      minority = percent_minority) %>%
    drop_na() %>%
    mutate(across(
      c(health, cesi, adversity, population_density, rurality, age, minority),
      ~ as.numeric(scale(.x))))

# M1 adjusts for baseline county characteristics and allows state intercepts to vary
  m1 <- lmerTest::lmer(
    health ~ cesi + population_density + rurality + age + minority +
      (1 | state),
    data = model_data,
    REML = FALSE)

# M2 adds county adversity score
  m2 <- lmerTest::lmer(
    health ~ cesi + adversity + population_density + rurality + age +
      minority + (1 | state),
    data = model_data,
    REML = FALSE)

  m1_coefficient <- coef(summary(m1))["cesi", ]
  m2_coefficient <- coef(summary(m2))["cesi", ]

  m1_variance <- as.data.frame(lme4::VarCorr(m1))
  m2_variance <- as.data.frame(lme4::VarCorr(m2))

  m1_state_variance <- m1_variance$vcov[m1_variance$grp == "state"]
  m1_residual_variance <- m1_variance$vcov[m1_variance$grp == "Residual"]
  m2_state_variance <- m2_variance$vcov[m2_variance$grp == "state"]
  m2_residual_variance <- m2_variance$vcov[m2_variance$grp == "Residual"]

  tibble(
    outcome_id = selected_outcome,
    cesi_measure = cesi_variable,
    model = c("MLM-M1", "MLM-M2"),
    beta = c(m1_coefficient[1], m2_coefficient[1]),
    standard_error = c(m1_coefficient[2], m2_coefficient[2]),
    degrees_freedom = c(m1_coefficient[3], m2_coefficient[3]),
    t = c(m1_coefficient[4], m2_coefficient[4]),
    p = c(m1_coefficient[5], m2_coefficient[5]),
    n = nrow(model_data),
    n_states = n_distinct(model_data$state),
    icc = c(
      m1_state_variance / (m1_state_variance + m1_residual_variance),
      m2_state_variance / (m2_state_variance + m2_residual_variance)),
    singular = c(
      lme4::isSingular(m1, tol = 1e-4),
      lme4::isSingular(m2, tol = 1e-4))) # check singularity for each model
}
```

```{r run-models}
# Run both models for practice and meaning cesi
outcome_ids <- unique(health_data$outcome_id)
cesi_variables <- c("practice_cesi", "meaning_cesi")

model_results_list <- list()
result_number <- 1

for (selected_outcome in outcome_ids) {
  for (cesi_variable in cesi_variables) {
    model_results_list[[result_number]] <- run_primary_models(
      health_data,
      selected_outcome,
      cesi_variable)

    result_number <- result_number + 1
  }
}

# Adjust for p values 
primary_results <- bind_rows(model_results_list) %>%
  group_by(model, cesi_measure) %>%
  mutate(p_holm = p.adjust(p, method = "holm")) %>%
  ungroup() %>%
  left_join(
    health %>%
      distinct(
        source,
        outcome_id,
        outcome,
        native_unit,
        higher_value_interpretation),
    by = "outcome_id") %>%
  mutate(
    cesi_measure = recode(
      cesi_measure,
      practice_cesi = "Practice CESI",
      meaning_cesi = "Meaning CESI")) %>%
  relocate(
    source,
    outcome_id,
    outcome,
    native_unit,
    higher_value_interpretation,
    cesi_measure,
    model) %>%
  arrange(cesi_measure, model, outcome)
primary_results

```

# Model checks

```{r model-checks}
model_checks <- primary_results %>%
  summarise(
    models = n(),
    minimum_n = min(n),
    maximum_n = max(n),
    minimum_states = min(n_states),
    maximum_states = max(n_states),
    singular_models = sum(singular))
model_checks

# Verify M1 and M2 used the same counties and number of states
same_sample_check <- primary_results %>%
  select(outcome_id, cesi_measure, model, n, n_states) %>%
  pivot_wider(
    names_from = model,
    values_from = c(n, n_states)) %>%
  mutate(
    same_n = `n_MLM-M1` == `n_MLM-M2`,
    same_number_of_states =
      `n_states_MLM-M1` == `n_states_MLM-M2`)

same_sample_check %>%
  filter(!same_n | !same_number_of_states)

```

# Export

```{r export}
output_dir <- "output/primary_health_models"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

analysis_data_dir <- "data/derived/health_analysis"
dir.create(analysis_data_dir, recursive = TRUE, showWarnings = FALSE)

write_csv(health_data, file.path(analysis_data_dir, "county_health_analysis_long.csv"))
write_csv(primary_results, file.path(output_dir, "primary_mlm_results.csv"))
write_csv(sample_coverage, file.path(output_dir, "primary_model_sample_coverage.csv"))
write_csv(model_checks, file.path(output_dir, "primary_model_checks.csv"))

```
