---
title: "CESI and County Health Sensitivity Analyses"
author: "Erin Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: readable
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))

knitr::opts_chunk$set(
  echo = TRUE,
  warning = TRUE,
  message = FALSE,
  error = FALSE)

library(dplyr)
library(tidyr)
library(readr)
library(stringr)
```

This script is for sensitivity analysis.

# Read in data

```{r read-data}
health_data <- read_csv(
  "data/derived/health_analysis/county_health_analysis_long.csv",
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE)

outcome_information <- health_data %>%
  distinct(
    source,
    outcome_id,
    outcome,
    native_unit,
    higher_value_interpretation)

outcome_ids <- unique(health_data$outcome_id)

cesi_variables <- c("practice_cesi", "meaning_cesi")

```

# Bivariate correlations

These correlations are the unadjusted county-level associations.

```{r correlations}
# Store result for outcome-CESI 
correlation_list <- list()
result_number <- 1

# Correlation
for (selected_outcome in outcome_ids) {
  for (cesi_variable in cesi_variables) {
    correlation_data <- health_data %>%
      filter(outcome_id == selected_outcome) %>%
      transmute(
        health = outcome_value,
        cesi = .data[[cesi_variable]]) %>%
      drop_na()

    correlation_test <- cor.test(
      correlation_data$cesi,
      correlation_data$health)

    correlation_list[[result_number]] <- tibble(
      outcome_id = selected_outcome,
      cesi_measure = cesi_variable,
      r = unname(correlation_test$estimate),
      p = correlation_test$p.value,
      n = nrow(correlation_data))

    result_number <- result_number + 1
  }
}

correlation_results <- bind_rows(correlation_list) %>%
  group_by(cesi_measure) %>%
  mutate(p_holm = p.adjust(p, method = "holm")) %>%
  ungroup()

```

# linear regressions and state fixed effects

- The linear regressions did not include state-level clustering. 
- The state fixed-effects model compares counties within the same state and uses standard errors clustered by state.

```{r ols-and-state-fixed-effects}
# Fit OLS and state fixed-effects models
geographic_list <- list()
result_number <- 1

for (selected_outcome in outcome_ids) {
  for (cesi_variable in cesi_variables) {
    model_data <- health_data %>%
      filter(outcome_id == selected_outcome) %>%
      transmute(
        state = factor(state),
        health = outcome_value,
        cesi = .data[[cesi_variable]],
        adversity = adversity_domain_balanced,
        population_density = log_population_density,
        rurality = rural_urban_code,
        age = county_age,
        minority = percent_minority) %>%
      drop_na() %>%
      mutate(across(
        c(health, cesi, adversity, population_density, rurality, age, minority),
        ~ as.numeric(scale(.x))))

# Fit linear regression
    ols <- lm(
      health ~ cesi + adversity + population_density + rurality + age +
        minority,
      data = model_data)

# Adding state fixed effect
    state_fe <- lm(
      health ~ cesi + adversity + population_density + rurality + age +
        minority + state,
      data = model_data)

    ols_coefficient <- coef(summary(ols))["cesi", ]

# Cluster the fixed-effects model's standard errors by state because counties in the same state 
    state_vcov <- sandwich::vcovCL(
      state_fe,
      cluster = ~ state,
      type = "HC1")

    state_beta <- coef(state_fe)["cesi"]
    state_se <- sqrt(state_vcov["cesi", "cesi"])
    state_t <- state_beta / state_se
    state_p <- 2 * pt(
      abs(state_t),
      df = n_distinct(model_data$state) - 1,
      lower.tail = FALSE)

    geographic_list[[result_number]] <- tibble(
      outcome_id = selected_outcome,
      cesi_measure = cesi_variable,
      model = c("OLS", "State fixed effects"),
      beta = c(ols_coefficient[1], state_beta),
      standard_error = c(ols_coefficient[2], state_se),
      p = c(ols_coefficient[4], state_p),
      n = nrow(model_data),
      n_states = n_distinct(model_data$state))

    result_number <- result_number + 1
  }
}

# Adjust for p values
geographic_results <- bind_rows(geographic_list) %>%
  group_by(model, cesi_measure) %>%
  mutate(p_holm = p.adjust(p, method = "holm")) %>%
  ungroup()

```

# Separate within-state and between-state associations

The county CESI score is separated into the state mean and each county's deviation from that mean. The within-state coefficient compares counties in the same state. The between-state coefficient compares state-average CESI levels.

```{r within-between}
within_between_list <- list()
result_number <- 1

for (selected_outcome in outcome_ids) {
  for (cesi_variable in cesi_variables) {
# Get the state means
    state_means <- health_data %>%
      distinct(county_fips, state, cesi = .data[[cesi_variable]]) %>%
      filter(!is.na(cesi)) %>%
      group_by(state) %>%
      summarise(cesi_state_mean = mean(cesi), .groups = "drop")

    model_data <- health_data %>%
      filter(outcome_id == selected_outcome) %>%
      transmute(
        county_fips,
        state,
        health = outcome_value,
        cesi = .data[[cesi_variable]],
        adversity = adversity_domain_balanced,
        population_density = log_population_density,
        rurality = rural_urban_code,
        age = county_age,
        minority = percent_minority) %>%
      left_join(state_means, by = "state") %>%
      mutate(
        cesi_within_state = cesi - cesi_state_mean,
        state = factor(state)) %>%
      drop_na() %>%
      mutate(across(
        c(
          health, cesi_within_state, cesi_state_mean, adversity,
          population_density, rurality, age, minority),
        ~ as.numeric(scale(.x))))

    model <- lmerTest::lmer(
      health ~ cesi_within_state + cesi_state_mean + adversity +
        population_density + rurality + age + minority + (1 | state), # state random intercept 
      data = model_data,
      REML = FALSE)

    coefficient_table <- coef(summary(model))

    within_between_list[[result_number]] <- tibble(
      outcome_id = selected_outcome,
      cesi_measure = cesi_variable,
      effect = c("Within state", "Between states"),
      beta = coefficient_table[
        c("cesi_within_state", "cesi_state_mean"), 1],
      standard_error = coefficient_table[
        c("cesi_within_state", "cesi_state_mean"), 2],
      p = coefficient_table[
        c("cesi_within_state", "cesi_state_mean"), 5],
      n = nrow(model_data),
      n_states = n_distinct(model_data$state))

    result_number <- result_number + 1
  }
}

# Adjust for p values
within_between_results <- bind_rows(within_between_list) %>%
  group_by(effect, cesi_measure) %>%
  mutate(p_holm = p.adjust(p, method = "holm")) %>%
  ungroup()

```

# Joint Practice and Meaning CESI model

This model estimates each CESI association while holding the other CESI measure constant.

```{r joint-model}
joint_list <- list()

for (i in seq_along(outcome_ids)) {
  selected_outcome <- outcome_ids[i]

  model_data <- health_data %>%
    filter(outcome_id == selected_outcome) %>%
    transmute(
      state = factor(state),
      health = outcome_value,
      practice = practice_cesi,
      meaning = meaning_cesi,
      adversity = adversity_domain_balanced,
      population_density = log_population_density,
      rurality = rural_urban_code,
      age = county_age,
      minority = percent_minority) %>%
    drop_na() %>%
    mutate(across(
      c(
        health, practice, meaning, adversity,
        population_density, rurality, age, minority),
      ~ as.numeric(scale(.x))))

# M2 with both CESI
  model <- lmerTest::lmer(
    health ~ practice + meaning + adversity + population_density +
      rurality + age + minority + (1 | state),
    data = model_data,
    REML = FALSE)

  coefficient_table <- coef(summary(model))

  joint_list[[i]] <- tibble(
    outcome_id = selected_outcome,
    cesi_measure = c("practice_cesi", "meaning_cesi"),
    beta = coefficient_table[c("practice", "meaning"), 1],
    standard_error = coefficient_table[c("practice", "meaning"), 2],
    p = coefficient_table[c("practice", "meaning"), 5],
    n = nrow(model_data),
    n_states = n_distinct(model_data$state))
}

# Adjust for p values
joint_results <- bind_rows(joint_list) %>%
  group_by(cesi_measure) %>%
  mutate(p_holm = p.adjust(p, method = "holm")) %>%
  ungroup()

```

# Four adversity domains

Seperately testing if use four adversity domains separately instead of combining them into one score yeild similar results. 

```{r four-domain-model}
domain_list <- list()
result_number <- 1

for (selected_outcome in outcome_ids) {
  for (cesi_variable in cesi_variables) {
    model_data <- health_data %>%
      filter(outcome_id == selected_outcome) %>%
      transmute(
        state = factor(state),
        health = outcome_value,
        cesi = .data[[cesi_variable]],
        socioeconomic = socioeconomic_z,
        material_hardship = material_hardship_z,
        healthcare_access = healthcare_access_z,
        preventive_resources = preventive_resources_z,
        population_density = log_population_density,
        rurality = rural_urban_code,
        age = county_age,
        minority = percent_minority) %>%
      drop_na() %>% #include counties with all four domains and covariates
      mutate(across(
        c(
          health, cesi, socioeconomic, material_hardship,
          healthcare_access, preventive_resources,
          population_density, rurality, age, minority),
        ~ as.numeric(scale(.x))))

# Model the four adversity domains separately
    model <- lmerTest::lmer(
      health ~ cesi + socioeconomic + material_hardship +
        healthcare_access + preventive_resources + population_density +
        rurality + age + minority + (1 | state),
      data = model_data,
      REML = FALSE)

    coefficient <- coef(summary(model))["cesi", ]

    domain_list[[result_number]] <- tibble(
      outcome_id = selected_outcome,
      cesi_measure = cesi_variable,
      beta = coefficient[1],
      standard_error = coefficient[2],
      p = coefficient[5],
      n = nrow(model_data),
      n_states = n_distinct(model_data$state))

    result_number <- result_number + 1
  }
}

# Adjust for p values
four_domain_results <- bind_rows(domain_list) %>%
  group_by(cesi_measure) %>%
  mutate(p_holm = p.adjust(p, method = "holm")) %>%
  ungroup()

```

# CESI by adversity interactions

These models test whether the CESI association changes across levels of county adversity.

```{r interaction-models}
interaction_list <- list()
result_number <- 1

for (selected_outcome in outcome_ids) {
  for (cesi_variable in cesi_variables) {
    model_data <- health_data %>%
      filter(outcome_id == selected_outcome) %>%
      transmute(
        state = factor(state),
        health = outcome_value,
        cesi = .data[[cesi_variable]],
        adversity = adversity_domain_balanced,
        population_density = log_population_density,
        rurality = rural_urban_code,
        age = county_age,
        minority = percent_minority) %>%
      drop_na() %>%
      mutate(across(
        c(health, cesi, adversity, population_density, rurality, age, minority),
        ~ as.numeric(scale(.x))))

# Include both main effects with their interaction and retain state clustering.
    model <- lmerTest::lmer(
      health ~ cesi * adversity + population_density + rurality + age +
        minority + (1 | state),
      data = model_data,
      REML = FALSE)

    coefficient <- coef(summary(model))["cesi:adversity", ]

    interaction_list[[result_number]] <- tibble(
      outcome_id = selected_outcome,
      cesi_measure = cesi_variable,
      beta = coefficient[1],
      standard_error = coefficient[2],
      p = coefficient[5],
      n = nrow(model_data),
      n_states = n_distinct(model_data$state))

    result_number <- result_number + 1
  }
}

# Adjust for p values
interaction_results <- bind_rows(interaction_list) %>%
  group_by(cesi_measure) %>%
  mutate(p_holm = p.adjust(p, method = "holm")) %>%
  ungroup()

```

# Export

```{r export}
# Add labels
correlation_results <- correlation_results %>%
  left_join(outcome_information, by = "outcome_id")

geographic_results <- geographic_results %>%
  left_join(outcome_information, by = "outcome_id")

within_between_results <- within_between_results %>%
  left_join(outcome_information, by = "outcome_id")

joint_results <- joint_results %>%
  left_join(outcome_information, by = "outcome_id")

four_domain_results <- four_domain_results %>%
  left_join(outcome_information, by = "outcome_id")

interaction_results <- interaction_results %>%
  left_join(outcome_information, by = "outcome_id")

# Export
output_dir <- "output/sensitivity_analyses"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

write_csv(correlation_results, file.path(output_dir, "correlations.csv"))
write_csv(geographic_results, file.path(output_dir, "ols_and_state_fixed_effects.csv"))
write_csv(within_between_results, file.path(output_dir, "within_between_state_results.csv"))
write_csv(joint_results, file.path(output_dir, "joint_cesi_mlm_results.csv"))
write_csv(four_domain_results, file.path(output_dir, "four_adversity_domain_results.csv"))
write_csv(interaction_results, file.path(output_dir, "cesi_adversity_interactions.csv"))

```
