---
title: "CESI and County Health Tables and Figures"
author: "Erin Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: readable
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))

knitr::opts_chunk$set(
  echo = TRUE,
  warning = TRUE,
  message = FALSE,
  error = FALSE)

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
```

This script is for creating figures. Practice CESI is shown in blue, Meaning CESI in purple, and regression estimates in black.

# Read results

```{r read-results}
primary_results <- read_csv(
  "output/primary_health_models/primary_mlm_results.csv",
  show_col_types = FALSE)

health_data <- read_csv(
  "data/derived/health_analysis/county_health_analysis_long.csv",
  col_types = cols(county_fips = col_character()),
  show_col_types = FALSE)

within_between_results <- read_csv(
  "output/sensitivity_analyses/within_between_state_results.csv",
  show_col_types = FALSE)

# File paths
output_dir <- "output/tables_and_figures"
figure_dir <- file.path(output_dir, "figures")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(figure_dir, recursive = TRUE, showWarnings = FALSE)

```

# Outcome categories

```{r categories}
outcome_categories <- primary_results %>%
  distinct(outcome) %>%
  mutate(
    main_category = case_when(
      outcome %in% c(
        "Alcohol-Impaired Driving Deaths", "Firearm Fatalities",
        "Homicide Mortality", "Injury Deaths",
        "Motor Vehicle Crash Deaths", "HIV Prevalence",
        "Sexually Transmitted Infections", "Low Birth Weight",
        "Teen Births", "Child Mortality", "Infant Mortality",
        "Life Expectancy", "Premature Death",
        "Premature Age-Adjusted Mortality", "Binge Drinking",
        "Current Cigarette Smoking", "Physical Inactivity",
        "Short Sleep Duration", "Suicide Mortality",
        "Drug Overdose Deaths") ~ "Behavioral Risks",
      outcome %in% c(
        "Hearing Disability", "Independent Living Disability",
        "Mobility Disability", "Self-care Disability",
        "Vision Disability", "Cognitive Disability", "Any Disability",
        "Depression", "Frequent Mental Distress",
        "Lack of Social/Emotional Support", "Social Isolation",
        "General Health", "Frequent Physical Distress") ~
        "Subjective Wellbeing",
      TRUE ~ "Chronic Health"),
    subcategory = case_when(
      outcome %in% c(
        "Alcohol-Impaired Driving Deaths", "Firearm Fatalities",
        "Homicide Mortality", "Injury Deaths",
        "Motor Vehicle Crash Deaths") ~ "Injury and Violence Mortality",
      outcome %in% c(
        "HIV Prevalence", "Sexually Transmitted Infections",
        "Low Birth Weight", "Teen Births") ~
        "Sexual and Reproductive Risks",
      outcome %in% c(
        "Child Mortality", "Infant Mortality", "Life Expectancy",
        "Premature Death", "Premature Age-Adjusted Mortality") ~
        "Premature Deaths and Reduced Longevity",
      outcome %in% c(
        "Binge Drinking", "Current Cigarette Smoking",
        "Physical Inactivity", "Short Sleep Duration") ~
        "Health-Risk Behaviors",
      outcome %in% c("Suicide Mortality", "Drug Overdose Deaths") ~
        "Behavioral-Health Mortality",
      outcome %in% c(
        "Hearing Disability", "Independent Living Disability",
        "Mobility Disability", "Self-care Disability",
        "Vision Disability", "Cognitive Disability", "Any Disability") ~
        "Disability",
      outcome %in% c("Depression", "Frequent Mental Distress") ~
        "Mental Health",
      outcome %in% c(
        "Lack of Social/Emotional Support", "Social Isolation") ~
        "Social Connection and Support",
      outcome %in% c("General Health", "Frequent Physical Distress") ~
        "Subjective Health",
      outcome == "All Teeth Lost" ~ "Oral Health",
      TRUE ~ "Chronic Conditions"),
    display_outcome = if_else(
      outcome == "General Health",
      "Poor General Health", # Rename for clarity
      outcome))

```

# Presentation table

Stars are based on Holm-adjusted p values: `* p < .05`, `** p < .01`, and `*** p < .001`.

```{r presentation-table}
# Apply significance stars to Holm-adjusted, rather than raw, p values.
presentation_table <- primary_results %>%
  mutate(
    stars = case_when(
      p_holm < .001 ~ "***",
      p_holm < .01 ~ "**",
      p_holm < .05 ~ "*",
      TRUE ~ ""),
    coefficient = paste0(sprintf("%.2f", beta), stars)) %>%
  select(outcome, cesi_measure, model, coefficient) %>%
  pivot_wider(
    names_from = c(cesi_measure, model),
    values_from = coefficient,
    names_glue = "{cesi_measure} {model}") %>%
  left_join(outcome_categories, by = "outcome") %>%
  arrange(
    factor(
      main_category,
      levels = c("Behavioral Risks", "Subjective Wellbeing", "Chronic Health")),
    subcategory,
    display_outcome) %>%
  select(
    `Main Category` = main_category,
    `Subcategory` = subcategory,
    `Outcome` = display_outcome,
    `Practice MLM-M1` = `Practice CESI MLM-M1`,
    `Practice MLM-M2` = `Practice CESI MLM-M2`,
    `Meaning MLM-M1` = `Meaning CESI MLM-M1`,
    `Meaning MLM-M2` = `Meaning CESI MLM-M2`)
presentation_table

write_csv(presentation_table, file.path(output_dir, "primary_mlm_presentation_table.csv"))

```

# Relationship between the two CESI measures

```{r cesi-correlation, fig.width=7, fig.height=6}
cesi_plot_data <- health_data %>%
  distinct(county_fips, practice_cesi, meaning_cesi) %>%
  drop_na(practice_cesi, meaning_cesi)

cesi_correlation <- cor.test(
  cesi_plot_data$practice_cesi,
  cesi_plot_data$meaning_cesi)

cesi_correlation_results <- tibble(
  n = nrow(cesi_plot_data),
  r = unname(cesi_correlation$estimate),
  p = cesi_correlation$p.value)

# Plot
cesi_plot <- ggplot(
  cesi_plot_data,
  aes(x = practice_cesi, y = meaning_cesi)) +
  geom_point(color = "#118AB2", alpha = .45, size = 1.3) +
  geom_smooth(method = "lm", color = "black", fill = "grey80") +
  annotate(
    "text",
    x = -Inf,
    y = Inf,
    hjust = -.1,
    vjust = 1.4,
    label = sprintf(
      "n = %s, r = %.2f",
      nrow(cesi_plot_data),
      unname(cesi_correlation$estimate))) +
  labs(
    x = "Practice-based CESI",
    y = "Meaning-based CESI") +
  theme_classic(base_size = 12)
cesi_plot

ggsave(file.path(figure_dir, "practice_meaning_cesi_correlation.png"), cesi_plot, width = 7, height = 6, dpi = 300)

write_csv(cesi_correlation_results, file.path(output_dir, "practice_meaning_cesi_correlation.csv"))

```

# Figures for M2

```{r coefficient-figures, fig.width=9, fig.height=12}
coefficient_data <- primary_results %>%
  filter(model == "MLM-M2") %>%
  mutate(
    ci_low = beta - 1.96 * standard_error,
    ci_high = beta + 1.96 * standard_error) %>%
  left_join(outcome_categories, by = "outcome")

 # Plot
practice_plot <- coefficient_data %>%
  filter(cesi_measure == "Practice CESI") %>%
  ggplot(aes(x = beta, y = reorder(display_outcome, beta))) +
  geom_vline(xintercept = 0, color = "grey70") +
  geom_errorbar(
    aes(xmin = ci_low, xmax = ci_high),
    width = 0,
    orientation = "y") +
  geom_point(color = "#118AB2", size = 2) +
  facet_grid(main_category ~ ., scales = "free_y", space = "free_y") +
  labs(x = "Standardized coefficient", y = NULL) +
  theme_classic(base_size = 11)

meaning_plot <- coefficient_data %>%
  filter(cesi_measure == "Meaning CESI") %>%
  ggplot(aes(x = beta, y = reorder(display_outcome, beta))) +
  geom_vline(xintercept = 0, color = "grey70") +
  geom_errorbar(
    aes(xmin = ci_low, xmax = ci_high),
    width = 0,
    orientation = "y") +
  geom_point(color = "#6A4C93", size = 2) +
  facet_grid(main_category ~ ., scales = "free_y", space = "free_y") +
  labs(x = "Standardized coefficient", y = NULL) +
  theme_classic(base_size = 11)

practice_plot
meaning_plot

ggsave(file.path(figure_dir, "practice_cesi_primary_coefficients.png"), practice_plot, width = 9, height = 12, dpi = 300)

ggsave(file.path(figure_dir, "meaning_cesi_primary_coefficients.png"), meaning_plot, width = 9, height = 12, dpi = 300)

```
